### 2021-09-22 chapter1/chapter.tex
remove denominator in V(x) operator

Stochastic fluid queues have found a variety of applications such as telecommunications (see \cite{anick1982} as a canonical application in this area), power systems \cite{hydro}, risk processes \cite{betal2005} and environmental modelling \cite{wurm2020}. Fluid queues are relatively well studied. Largely, the analysis of fluid queues falls into two categories, matrix-analytic methods \cite{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}, and differential equation-based methods \cite{anick1982,kk1995,beanetal2019}. %For example, Ramaswami CITE, analysed fluid queues by mapping them to a quasi-birth-and-death process (QBD), after which they applied known matrix-analytic methods for QBDs to compute quantities of interest. Anick Mitra Sondhi CITE, analysed fluid queues using a more direct differential equation-based method. Since Ramaswami's CITE initial work, there has been significant developments in the analysis of fluid queues CITE and related algorithms CITE.

More recently, Bean and O'Reilly \cite{bo2014} extended fluid queues to \emph{so-called} stochastic fluid-fluid queues. In a fluid-fluid queue there is a second level process, \(\{Y(t)\}_{t\geq0}\) which is itself driven by a fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\). The analysis, Bean and O'Reilly \cite{bo2014}, is in principal similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\). For practical computation of the results \cite{bo2014} a matrix-discretisation of the infinitesimal generator of the fluid queue can be used. To this end, to date, two possible discretisation have been suggested. Taking a differential equations-based approach, Bean \emph{et al.}~\cite{beanetal2019} use the discontinuous Galerkin (DG) method to discretise this operator, while Bean and O'Reilly \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process. Both approaches are insightful and offer different tools and perspectives with which to analyse the resulting approximations. It turns out that the latter approach is a sub-class of the former; the QBD can be viewed as the simplest DG scheme where the operator is projected onto a basis of piecewise constant functions.

@@

In the context of approximating fluid queues, one advantage of the QBD discretisation and, equivalently, DG schemes with constant basis functions, is that they guarantee probabilities computed from the approximation are positive \cite[Section 3.3]{koltai2011}, see also \cite{nodalDGBook} and references therein. One justification for the positivity preserving property is from the interpretation of the discretisation as a stochastic process thereby ensuring positivity. For higher order DG schemes there is no such interpretation. Moreover, higher-order DG approximation schemes may produce negative, or highly oscillatory solutions, particularly when discontinuities or steep gradients are present. Methods to navigate the problem of negative or highly oscillatory solutions have been developed, such as slope limiters, and filtering (see \cite[Section 6.5]{nodalDGBook} and references therein). Slope limiting effectively alters the discretised operator in regions where oscillations are detected and lowers the order of the approximation in theses regions. Filtering is a post-hoc method which looks to recover an accurate solution, given an oscillatory approximation.

>>>>

Stochastic fluid queues have found a variety of applications such as telecommunications (see \cite{anick1982} as a canonical application in this area), power systems \citep{hydro}, risk processes \citep{betal2005} and environmental modelling \citep{wurm2020}. Fluid queues are relatively well studied. Largely, the analysis of fluid queues falls into two categories, matrix-analytic methods \citep{ajr2005,ar2003,ar2004,bean2005b,bean2005,bot08,bean2009,dasilva2005,latouche2018}, and differential equation-based methods \citep{anick1982,kk1995,beanetal2019}. %For example, Ramaswami CITE, analysed fluid queues by mapping them to a quasi-birth-and-death process (QBD), after which they applied known matrix-analytic methods for QBDs to compute quantities of interest. Anick Mitra Sondhi CITE, analysed fluid queues using a more direct differential equation-based method. Since Ramaswami's CITE initial work, there has been significant developments in the analysis of fluid queues CITE and related algorithms CITE.

More recently, \cite{bo2014} extended fluid queues to \emph{so-called} stochastic fluid-fluid queues. In a fluid-fluid queue there is a second level process, \(\{Y(t)\}_{t\geq0}\) which is itself driven by a fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\). The analysis,  \citep{bo2014}, is in principal similar to the matrix-analytic methods of \citep{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\). For practical computation of the results of \cite{bo2014}, a matrix-discretisation of the infinitesimal generator of the fluid queue can be used. To this end, to date, two possible discretisation have been suggested. Taking a differential equations-based approach, \cite{beanetal2019} use the discontinuous Galerkin (DG) method to discretise this operator, while \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process. Both approaches are insightful and offer different tools and perspectives with which to analyse the resulting approximations. It turns out that the latter approach is a sub-class of the former; the QBD can be viewed as the simplest DG scheme where the operator is projected onto a basis of piecewise constant functions.

@@

In the context of approximating fluid queues, one advantage of the QBD discretisation and, equivalently, DG schemes with constant basis functions, is that they guarantee probabilities computed from the approximation are positive \citep[Section 3.3]{koltai2011}, see also \citep{nodalDGBook} and references therein. One justification for the positivity preserving property is from the interpretation of the discretisation as a stochastic process thereby ensuring positivity. For higher order DG schemes there is no such interpretation. Moreover, higher-order DG approximation schemes may produce negative, or highly oscillatory solutions, particularly when discontinuities or steep gradients are present. Methods to navigate the problem of negative or highly oscillatory solutions have been developed, such as slope limiters, and filtering (see \citep[Section 6.5]{nodalDGBook} and references therein). Slope limiting effectively alters the discretised operator in regions where oscillations are detected and lowers the order of the approximation in theses regions. Filtering is a post-hoc method which looks to recover an accurate solution, given an oscillatory approximation.
